##Loading and preprocessing the data
###Load the data
```{r readData}
        activity<- read.csv('C:/Users/opeoluwa.osunkoya/Documents/data/repdata-data-activity/activity.csv')
```

###Convert the date variable to the date class
```{r PreProcess}
        class(activity$date)
        activity$date<- as.Date(activity$date)
        class(activity$date)
```
##What is the mean total number of steps taken per day?
###Calculate the total number of steps taken each day and plot an histogram
```{r}
        hist(totalSteps<- tapply(activity$steps, activity$date, FUN= sum, na.rm=TRUE))
```
### Calculate and report the mean and median total number of steps taken per day i.e.meanTNSTD and medianTNSTD
#### Mean
```{r}
        meanTNSTD <- tapply(activity$steps, activity$date, FUN= mean, na.rm=TRUE)
        print(as.table(meanTNSTD))
```
#### Median
```{r}
        medianTNSTD<- tapply(activity$steps, activity$date, FUN= median, na.rm=TRUE)
        print(as.table(medianTNSTD))
```
##What is the average daily activity pattern using 5 minute plots
-Time series plot 
```{r}
plot(type= "l",tapply(activity$steps, activity$interval, mean, na.rm=TRUE))
```
-To Find which 5-minute interval contains the maximum number of steps
```{r}
        activitymean<- tapply(activity$steps, activity$interval, mean, na.rm=TRUE)
        activitymean[activitymean==max(activitymean)]
```
##Inputting missing values
1. Total number of missing values in the dataset
```{r}
        table(is.na(activity))
```
2. Fill in the missing values
        -I'll use the mean for that 5 minute interval

3. Create a new data set equal to the original dataset but with the missing data filled in
        - First create a function to fill in the missing value
```{r ReplaceNAFunction }
                replaceNA<- function(IncompleteDataSet, DataSetWithAverages)
        {
                i<- dim(IncompleteDataSet)[1]
                for(a in 1:i)
                {
                        if(is.na(IncompleteDataSet$steps[i]))
                        {
                                IncompleteDataSet$steps[i]<- DataSetWithAverages[DataSetWithAverages= IncompleteDataSet$interval[i]]
                        }
                        a<-a+1
                }
                return(IncompleteDataSet)
        }
```
        -Then apply the function to the dataset
```{r FillInEmptyDataset}
        completeDataset<- replaceNA(activity, activitymean)
```
4. Compare the mean and median of the complete DataSet to that of the incomplete data set
        -Create the histogram
```{r}
        completeDataSetHistData<- tapply(completeDataset$steps, completeDataset$date, sum, na.rm=TRUE)
        hist(completeDataSetHistData)
```
        -The New Mean
```{r}
        meanCompleteDataSet <- tapply(completeDataset$steps, completeDataset$date, FUN= mean, na.rm=TRUE)
```
                -To compare with the existing mean, use the summary function. You'll find that they're the same.
```{r}
        summary(meanCompleteDataSet)
        summary(meanTNSTD)
```
                
        -The New Median. You'll find that they're the same.
```{r}
        medianCompleteDataSet<- tapply(completeDataset$steps, completeDataset$date, FUN= median, na.rm=TRUE)
```
                - To compare with the existing median
```{r}
        summary(medianCompleteDataSet)
        summary(medianTNSTD)
```
##Are there differences in activity patterns between weekdays and weekends?
1. Create a new function to denote whether the day is a weekday or a weekend in a new variable
```{r}
        CheckIfWeekend<- function(Dataset)
{
        Dataset$weekdayStat<- as.character(weekdays(Dataset$date))
        Dataset$weekdayStat[Dataset$weekdayStat=="Saturday"]<- "Weekend"
        Dataset$weekdayStat[Dataset$weekdayStat=="Sunday"]<- "Weekend"
        Dataset$weekdayStat[Dataset$weekdayStat!="Weekend"]<- "Weekday"
        return(Dataset)
}
```
        -Call the function
```{r}
completeDataset<- CheckIfWeekend(completeDataset)
```
2. Make a panel plot across all weekdays or weekend days
- Create a set for the weekdays
```{r}
        weekdaySet<- subset(completeDataset, subset= (completeDataset$weekdayStat=="Weekday"))
```
- Create a set for the weekends
```{r}
        weekendSet<- subset(completeDataset, subset= (completeDataset$weekdayStat=="Weekend"))
```       
-Plot for the weekdays
```{r}
        par(mfcol=c(2,1))
        WeekdayHistData<- tapply(weekdaySet$steps, weekdaySet$interval, sum, na.rm=TRUE)
        WeekendHistData<- tapply(weekendSet$steps, weekendSet$interval, sum, na.rm=TRUE)        
        plot(WeekdayHistData, type='l')
        plot(WeekendHistData, type='l')
```